---
title: "R Notebook"
output: html_notebook
---


```{r}
set.seed(1680)

library(dplyr)
library(ISLR)
library(cluster)
library(Rtsne)
library(ggplot2)
```



```{r}
getwd()
list.files()

rt <- read.csv("raw_trips_prepared_python.csv")
rc <- read.csv("raw_trips_transform_3.csv")
```


```{r}
# look at structure
str(rc)

# head
head(rc)
```

```{r}
# set seed
set.seed(1200)

```


```{r}
# clean data
rc$seats <- factor(rc$seats)
rc$dep_month <- factor(rc$dep_month)
rc$dep_dow <- factor(rc$dep_dow)
rc$children <- ifelse(rc$children == 0, 'False', 'True')
rc$children <- factor(rc$children)
rc$car <- ifelse(rc$car == 0, 'False', 'True')
rc$car <- factor(rc$car)
rc$hotel <- ifelse(rc$hotel == 0, 'False', 'True')
rc$hotel <- factor(rc$hotel)
rc$insurance <- ifelse(rc$insurance == 0, 'False', 'True')
rc$insurance <- factor(rc$insurance)

```

```{r}
head(rc)
```

```{r}
rc_vars <- rc %>%
  select(id, weekend_trip, Z_age, travel_companion)
```


# calculate distance
Defining (dis)similarity between observations
Euclidean is only for continuous variables, and is not applicable for mixed variables.
Gower distance can handle mixed data types.
Gower distance: for each var type, a particular metric that works well for that type is used and scaled to fall between 0 and 1 (e.x. - Manhattan for quantiative intervals, rank & manhattan for ordinal, dice coefficietn for nominal a.k.a., categorical vars)


```{r}
gower_dist_rc <- daisy(rc_vars[,-1],
                       metric = "gower")

summary(gower_dist_rc)
```

```{r}
gower_mat_rc <- as.matrix(gower_dist_rc)
```

```{r}
# find best number of k clusters
sil_width_rc <- c(NA)
for(i in 2:5) {
  pam_fit <- pam(gower_mat_rc,
                 diss = T,
                 k = i)
  sil_width_rc[i] <- pam_fit$silinfo$avg.width
}

# plot silhouette
qplot(x = c(1:5), y = sil_width_rc, geom = "point")
```



```{r}
pam_fit_rc <- pam(gower_dist_rc, diss = T,
                  k = 4)

pam_results_rc <- rc_vars %>%
  select(-id) %>%
  mutate(cluster = pam_fit_rc$clustering) %>%
  group_by(cluster) %>%
  do(the_summary = summary(.))

pam_results_rc$the_summary
```


```{r}
tsne_obj_rc <- Rtsne(gower_dist_rc, is_distance = T)

tsne_data_rc <- tsne_obj_rc$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit_rc$clustering),
         name = rc_vars$id)

ggplot(aes(x = X, y = Y), data = tsne_data_rc) +
  geom_point(aes(color = cluster))
```




```{r}
clu <- tsne_data_rc[,'cluster']
```



```{r}
rc <- rc %>%
  mutate(cluster = clu)
```

```{r}
head(rc)
names(rc)
```

```{r}
ggplot(rc) + 
  geom_bar(aes(cluster, y = ..count../sum(..count..),
               fill = travel_companion), position = 'dodge')
```

```{r}
ggplot(rc,aes(cluster, age)) +
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', color = I('red'))
```


```{r}
ggplot(rc) + 
  geom_bar(aes(cluster, fill = long_holiday), position = 'dodge')
```


```{r}
ggplot(rc) +
  geom_boxplot(aes(cluster, Z_duration))
```

```{r}

  ggplot(rc) + 
  geom_bar(aes(cluster, fill = weekend_trip), position = 'dodge')
```


```{r}
ggplot(rc) +
  geom_bar(aes(cluster, fill = gender), position = 'dodge')
```

```{r}
ggplot(rc) +
  geom_bar(aes(cluster, fill = children), position = 'dodge')
```

```{r}
ggplot(rc) +
  geom_freqpoly(aes(purchase_lag,..density.., color = cluster)) +
  geom_freqpoly(aes(purchase_lag, ..density..))

ggplot(rc, aes(x = cluster, y =purchase_lag)) +
  geom_boxplot() +
  stat_summary(fun.y = mean, color = I('red'), geom = 'point')
```



```{r}
ggplot(rc) +
  geom_bar(aes(cluster, fill = customer_type), position = 'dodge')
```

```{r}
ggplot(rc) +
  geom_bar(aes(x = cluster, fill = car), position = 'dodge')
```

```{r}
ggplot(rc) +
  geom_bar(aes(x = cluster, fill = hotel), position = 'dodge')
```

```{r}
ggplot(rc) +
  geom_bar(aes(x = cluster, fill = insurance), position = 'dodge')
```

```{r}
grid.arrange(
rc %>%
  filter(cluster == 1) %>%
  ggplot() +
  geom_bar(aes(x = dep_month, y= ..count../sum(..count..))),

rc %>%
  filter(cluster == 1) %>%
  ggplot() +
  geom_bar(aes(x = dep_dow, y= ..count../sum(..count..))),
ncol = 2)
```

```{r}
grid.arrange(
rc %>%
  filter(cluster == 2) %>%
  ggplot() +
  geom_bar(aes(x = dep_month, y= ..count../sum(..count..))),

rc %>%
  filter(cluster == 2) %>%
  ggplot() +
  geom_bar(aes(x = dep_dow, y= ..count../sum(..count..))),
ncol = 2)
```

```{r}
grid.arrange(
rc %>%
  filter(cluster == 3) %>%
  ggplot() +
  geom_bar(aes(x = dep_month, y= ..count../sum(..count..))),

rc %>%
  filter(cluster == 3) %>%
  ggplot() +
  geom_bar(aes(x = dep_dow, y= ..count../sum(..count..))),
ncol = 2)
```

```{r}
grid.arrange(
rc %>%
  filter(cluster == 4) %>%
  ggplot() +
  geom_bar(aes(x = dep_month, y= ..count../sum(..count..))),

rc %>%
  filter(cluster == 4) %>%
  ggplot() +
  geom_bar(aes(x = dep_dow, y= ..count../sum(..count..))),
ncol = 2)
```

